detect patches

### 2024-01-23 scripts/classify_patches.Rmd
a few patch metrics before meeting

>>>>

```{r Patch metics to predict presence}

# euclidean nearest neighbor distance (enn)
## (modification) log(x), sqrt(x)
### Expectation: the number of colonized predicted patches decrease with increasing distance.
#### why: long-distance dispersal is limited from colonized to suitable patches. 
##### test: logistic regression, independent: distance, dependent: colonization (0|1)

sqrt(c( 1000, 150))
```

```{r Patch metrics to predict abundance}
library(landscapemetrics)

# Patch level metrics. 
print(
  list_lsm(level = "patch"), 
  n = 12)

# Core Area Index (cai)  
## (modification) arithmetic mean of core area cells
### Expectation: higher cai higher abundance
#### why: better within patch distribution, and a persistent 'source' to 'sink' peripheral cells. 
##### test: glmm?(% cover ~ cai, correlation = nlme::corExp(Long ~ Lat, nugget=T), family = quasibinomial)

# perimeter-area ratio (para)
## (modification) none
### Expectation: lower para higher abundance
#### why: areas with 'large edges' will suffer dispersal limitations throughout them, as many seeds will go to unsuitable habitat.
##### test: glmm?(% cover ~ para, correlation = nlme::corExp(Long ~ Lat, nugget=T), family = quasibinomial)

# fractal dimension index (frac)
## (modification) none
### Expectation: lower fractal dimensions (less complex patch shapes) have higher abundances. 
#### why: components of patches with complex shapes may suffer dispersal limitation. 
##### test: glmm?(% cover ~ frac, correlation = nlme::corExp(Long ~ Lat, nugget=T), family = quasibinomial)

## to all models a term for the previous 4 years SPEI may be incorporated.

```
